---
title: "Master Thesis Code"
author: "Nina van Gerwen (1860852)"
date: "2022-10-28"
output: html_document
---

## Functions for IRF calculations

These are necessary in data generation as we will vary the data generating model.

```{r}
source("Functions/model-calculations.R")
source("Functions/data-generation.R")
source("Functions/missingness.R")
source("Functions/data-aggregation.R")
source("Functions/lr-test-calculation.R")
source("Functions/chi-square-diff-test.R")
source("Functions/pearson-chi-sq-test.R")
source("Functions/start.values.R")
```

## Functions for model fit calculations

To fit the 1PL model, we can use the rasch() function.
To fit the 2PL model, we can use the ltm() function.
To fit the 3PL model, we can use the tpm() function.

## Empirical alpha analysis

```{r Final first run}
library(ltm)
library(dplyr)
library(mirt)
## This is for empirical alpha estimation
test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 300, 500, 1000, 1500))

n_sim <- 300

## We vary three factors, and for the simulation study, we 
## cross-examine them through a nested for loop

alpha_results <- data.frame(test_length = NA, sample_size = NA, LR2 = NA,
                      LR3 = NA, LR4 = NA, Chisq = NA, P_Chisq = NA)

set.seed(1248)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
      
      prop_1 <- rep(NA, n_sim)
      prop_2 <- rep(NA, n_sim)
      prop_3 <- rep(NA, n_sim)
      prop_4 <- rep(NA, n_sim)
      prop_5 <- rep(NA, n_sim)
      
      temp_n <- as.numeric(as.character(b))
      temp_k <- as.numeric(as.character(a))
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){
        ## Then, for every condition, generate data according to the
        ## current condition

          temp_data <- data.gen(n = temp_n, k = temp_k, 
                                model = "2PL")
          
          temp_model <- ltm(temp_data ~ z1, IRT.param = TRUE, start.val = start.vals)
          
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 2, start.val = start.vals), 
                              df = 2 * temp_k)
 
          p_value_2 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 3, start.val = start.vals),
                              df = 2 * (2 *temp_k))
        
          p_value_3 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 4, start.val = start.vals), 
                              df = 3 * (2 *temp_k))
          
          p_value_4 <- equal.mirt.diff.test(l_0 = l_0, dataset = temp_data, k = temp_k)
          
          if(temp_k == 20){
          p_value_5 <- 1
          } else {
          p_value_5 <- pearson.test(model = temp_model, k = temp_k)
          }
          
          
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)
          prop_2[i] <- ifelse(p_value_2 < .05, 1, 0)
          prop_3[i] <- ifelse(p_value_3 < .05, 1, 0)
          prop_4[i] <- ifelse(p_value_4 < .05, 1, 0)
          prop_5[i] <- ifelse(p_value_5 < .05, 1, 0)
        }
      
      alpha_results <- rbind(alpha_results, c(a, b, mean(prop_1), mean(prop_2), 
                                  mean(prop_3), mean(prop_4), mean(prop_5)))
  }
  
}

alpha_results <- alpha_results[-1 ,]
rownames(alpha_results) <- NULL

round(alpha_results, digits = 3)
```


## Power analysis

```{r Final second run}
## This is for power estimation
set.seed(1248)

n_sim <- 300

## We vary three factors, and for the simulation study, we 
## cross-examine them through a nested for loop

## TUPLES

power_results <- data.frame(test_length = NA, sample_size = NA, LR2 = NA,
                      LR3 = NA, LR4 = NA, Chisq = NA, P_Chisq = NA)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
      
      prop_1 <- rep(NA, n_sim)
      prop_2 <- rep(NA, n_sim)
      prop_3 <- rep(NA, n_sim)
      prop_4 <- rep(NA, n_sim)
      prop_5 <- rep(NA, n_sim)
      
      temp_n <- as.numeric(as.character(b))
      temp_k <- as.numeric(as.character(a))
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){
        ## Then, for every condition, generate data according to the
        ## current condition

          temp_data <- data.gen(n = temp_n, 
                                k = temp_k, 
                                model = "3PL")
          
          temp_model <- ltm(temp_data ~ z1, IRT.param = TRUE, start.val = start.vals)
          
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 2, start.val = start.vals), 
                              df = 2 * temp_k)
 
          p_value_2 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 3, start.val = start.vals),
                              df = 2 * (2 *temp_k))
        
          p_value_3 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 4, start.val = start.vals), 
                              df = 3 * (2 *temp_k))
          
          p_value_4 <- equal.mirt.diff.test(l_0 = l_0, dataset = temp_data, k = temp_k)
          
          if(temp_k == 20){
          p_value_5 <- 1
          } else {
          p_value_5 <- pearson.test(model = temp_model, k = temp_k)
          }
        
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)
          prop_2[i] <- ifelse(p_value_2 < .05, 1, 0)
          prop_3[i] <- ifelse(p_value_3 < .05, 1, 0)
          prop_4[i] <- ifelse(p_value_4 < .05, 1, 0)
          prop_5[i] <- ifelse(p_value_5 < .05, 1, 0)
        
          }
      
      power_results <- rbind(power_results, c(a, b, mean(prop_1), mean(prop_2), 
                                  mean(prop_3), mean(prop_4), mean(prop_5)))

  }
  
}

power_results <- power_results[-1 ,]
rownames(power_results) <- NULL

power_results
```

### Testing with 3PL

```{r}
mirt.diff.test <- function(l_0, dataset, k){
  l_a <- dataset %>% 
            as.data.frame(.) %>%
            mirt(data = ., model = 1, itemtype = "3PL", optimizer = "nlminb", 
                 TOL = .0001, control = list(maxit = 1000),
                 verbose = TRUE, constrain = list(c(3, 7, 11, 15, 19))) %>% logLik(.)
  
  chisqvalue <- 2 * (l_a - l_0)
  
  p.value <- 1 - pchisq(chisqvalue, df = 1)
  
  return(p.value)
}

set.seed(1248)
p_value_4 <- rep(NA, 100)
for(i in 1:100){
          temp_data <- temp.data.gen(n = 1000, 
                                k = 5, 
                                model = "2PL")
          
          temp_model <- ltm(temp_data ~ z1, IRT.param = TRUE)
          
          l_0 <- temp_model$log.Lik
          
          p_value_4[i] <- mirt.diff.test(l_0 = l_0, dataset = temp_data, k = 5)
          
          print(p_value_4[i])
}

sum(ifelse(p_value_4 < .05, 1, 0))

set.seed(1248)
for(i in 1:100){
          temp_data <- temp.data.gen(n = 1000, 
                                k = 5, 
                                model = "3PL")
          
          temp_model <- ltm(temp_data ~ z1, IRT.param = TRUE)
          
          l_0 <- temp_model$log.Lik
          
          p_value_4[i] <- mirt.diff.test(l_0 = l_0, dataset = temp_data, k = 5)
          
          print(p_value_4[i])
}

sum(ifelse(p_value_4 < .05, 1, 0))
```
    


## Testing with looping

```{r}
test <- data.frame(TL = NA, SS = NA)
for(a in levels(test_length)){ 
  for(b in levels(sample_size)){
    
    temp_n <- as.numeric(as.character(b)) 
    temp_k <- as.numeric(as.character(a))
  for(i in 1:10){
    
    temp_data <- data.gen(n = temp_n, k = temp_k, "2PL")
    
    print(dim(temp_data))
    }
    
    test <- rbind(c(dim(temp_data)), test)

  }
  
}
test
```


## Testing with 1PL


```{r Testing with 1PL - 2PL}
set.seed(1248)

n_sim <- 300

## We vary three factors, and for the simulation study, we 
## cross-examine them through a nested for loop

## TUPLES

one_pl_power_results <- data.frame(test_length = NA, sample_size = NA, LR2 = NA,
                      LR3 = NA, LR4 = NA, Chisq = NA, P_Chisq = NA)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
      
      prop_1 <- rep(NA, n_sim)
      prop_2 <- rep(NA, n_sim)
      prop_3 <- rep(NA, n_sim)
      prop_4 <- rep(NA, n_sim)
      prop_5 <- rep(NA, n_sim)
      
      temp_n <- as.numeric(as.character(b))
      temp_k <- as.numeric(as.character(a))
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){
        ## Then, for every condition, generate data according to the
        ## current condition

          temp_data <- data.gen(n = temp_n, 
                                k = temp_k, 
                                model = "2PL")
          
          temp_model <- rasch(temp_data)
          
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 2), 
                              df = 2 * temp_k)
 
          p_value_2 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 3),
                              df = 2 * (2 *temp_k))
        
          p_value_3 <- 1 - pchisq(q = LR.test(l_0 = l_0, data = temp_data, g = 4), 
                              df = 3 * (2 *temp_k))
          
          two_pl_model <- ltm(temp_data ~ z1, IRT.param = TRUE, start.val = start.vals)
          
          l_a <- two_pl_model$log.Lik
          p_value_4 <- 1 - pchisq(q = 2*(l_a - l_0), df = temp_k)
          
          if(temp_k == 20){
          p_value_5 <- 1
          } else {
          p_value_5 <- pearson.test(model = temp_model, k = temp_k)
          }
        
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)
          prop_2[i] <- ifelse(p_value_2 < .05, 1, 0)
          prop_3[i] <- ifelse(p_value_3 < .05, 1, 0)
          prop_4[i] <- ifelse(p_value_4 < .05, 1, 0)
          prop_5[i] <- ifelse(p_value_5 < .05, 1, 0)
        
          }
      
      one_pl_power_results <- rbind(one_pl_power_results, c(a, b, mean(prop_1), mean(prop_2), 
                                  mean(prop_3), mean(prop_4), mean(prop_5)))

  }
  
}

one_pl_power_results <- one_pl_power_results[-1 ,]
rownames(one_pl_power_results) <- NULL

one_pl_power_results
```


