---
title: "Simulation Study II: Detecting Group Differences"
author: "Nina van Gerwen (1860852)"
date: "24-03-2023"
output: html_document
---

## 1. Preparation

To run the simulations, we load functions that we created ourselves. For this, we use source() and choose the appropriate files. To 
find out how the functions work exactly, please look at the function files directly.

```{r Function loading }
source("Functions/model-calculations.R")
source("Functions/start.values.R")
source("Functions/multigroup-data-gen.R")
source("Functions/multigroup-lr-test-calculation.R")
```

## 2. Power analysis for differences in item parameters alpha and beta

### 2.1. 2PL Condition

```{r Item Param Power 2PL}
library(ltm)      ## to fit the 2PL
library(dplyr)    ## for some tidying and the pipe
library(tidyr)    ## for more tidying

## Set the factor levels and number of simulations for the study
test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 300, 500, 1000, 1500))
n_groups <- as.factor(c(2, 3, 4))

n_sim <- 300

## We vary three factors, which we cross-examine them through a nested for loop

## First, we want a dataset to hold the results
IPP_a_b_2PL <- data.frame(test_length = NA, sample_size = NA, G = NA, MG.LR = NA)

## Set a seed for reproducibility
set.seed(1248)

## Then, start the nested for loop with all factors
for(a in levels(test_length)){

  for(b in levels(sample_size)){
    
    for(d in levels(n_groups)){
      
      ## In the end, we want a vector which will contain 0's and 1's 
      ## for each time the hypothesis is respectively not rejected or rejected
      ## so, we first create space of size equal to the number of simulations
      prop_1 <- rep(NA, n_sim)
      
      ## Temporarily save all the current levels of the factors as numerics
      temp_k <- as.numeric(as.character(a))
      temp_g <- as.numeric(as.character(d))
      temp_n <- round(as.numeric(as.character(b))/temp_g)
      
      ## To improve the convergence of the models we use, we work with 
      ## the start values function
      start.vals <- start.values(temp_k)
      
        ## Now, for each condition:
        for(i in 1:n_sim){

          ## Generate data according to current parameters
          temp_data <- multigroup.data.gen(n = temp_n, k = temp_k, g = temp_g, 
                                           method = "param", model = "2PL")
            
          ## Fit the 2PL model
          temp_model <- ltm(temp_data[ , 1:temp_k] ~ z1, IRT.param = TRUE)
          
          ## Get the likelihood of the fitted 2PL model
          l_0 <- temp_model$log.Lik
        
          ## Calculate the p-value of our MG LR test
          p_value_1 <- 1 - pchisq(q = MG.LR.test(l_0 = l_0, data = temp_data, g = temp_g, 
                                                 k = temp_k, start.val = start.vals), 
                              df = (temp_g - 1) * 2 * temp_k)
 
          ## Check whether the null hypothesis is rejected or not
          ## and put this in the ith value of the earlier created vector
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)

        }
        
        ## Now, to get the rejection rate, we use rbind and get the mean of 
        ## the earlier created vector as the rejection rate
        IPP_a_b_2PL <- rbind(IPP_a_b_2PL, c(a, b, d, mean(prop_1)))
      }
      

  }
  
}

## Some pivoting and subsetting in order to get a nicer overview of the results
IPP_a_b_2PL <- IPP_a_b_2PL %>% 
  pivot_wider(., names_from = G, values_from = MG.LR) %>%
  .[-1, -3]

## Resetting the rownames and colnames of the results 
rownames(IPP_a_b_2PL) <- NULL
colnames(IPP_a_b_2PL) <- c("test_length", "sample_size", "LR2", "LR3", "LR4")

## Calling the results
IPP_a_b_2PL
```

### 2.2. 3PL Condition

```{r Item Param Power 3PL}
library(ltm)
library(dplyr)
library(mirt)
library(tidyr)

test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 300, 500, 1000, 1500))
n_groups <- as.factor(c(2, 3, 4))

n_sim <- 300

IPP_a_b_3PL <- data.frame(test_length = NA, sample_size = NA, G = NA, MG.LR = NA)

set.seed(1248)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
    
    for(d in levels(n_groups)){
      
      prop_1 <- rep(NA, n_sim)
      
      temp_k <- as.numeric(as.character(a))
      temp_g <- as.numeric(as.character(d))
      temp_n <- round(as.numeric(as.character(b))/temp_g)
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){
          
          ## For this study, everything is the same, except that data is
          ## generated according to the 3PL model
          temp_data <- multigroup.data.gen(n = temp_n, k = temp_k, g = temp_g, 
                                           method = "param", model = "3PL")
            
          temp_model <- ltm(temp_data[ , 1:temp_k] ~ z1, IRT.param = TRUE)
          
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = MG.LR.test(l_0 = l_0, data = temp_data, g = temp_g, 
                                                 k = temp_k, start.val = start.vals), 
                              df = (temp_g - 1) * 2 * temp_k)
 
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)

        }
      
        IPP_a_b_3PL <- rbind(IPP_a_b_3PL, c(a, b, d, mean(prop_1)))
      }
      

  }
  
}

IPP_a_b_3PL <- IPP_a_b_3PL %>%
  pivot_wider(., names_from = G, values_from = MG.LR) %>%
  .[-1, -3]

rownames(IPP_a_b_3PL) <- NULL
colnames(IPP_a_b_3PL) <- c("test_length", "sample_size", "LR2", "LR3", "LR4")

IPP_a_b_3PL
```


## 3. Power analysis for differences in theta

### 3.1 2PL Condition

```{r Theta Power 2PL }
library(ltm)
library(dplyr)
library(tidyr)

test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 300, 500, 1000, 1500))
n_groups <- as.factor(c(2, 3, 4))

n_sim <- 300

TP_2PL <- data.frame(test_length = NA, sample_size = NA, G = NA, MG.LR = NA)

set.seed(1248)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
    
    for(d in levels(n_groups)){
      
      prop_1 <- rep(NA, n_sim)
      
      temp_k <- as.numeric(as.character(a))
      temp_g <- as.numeric(as.character(d))
      temp_n <- round(as.numeric(as.character(b))/temp_g)
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){

          ## For this, everything is again the same as in the first 
          ## annotated study, except that we now simulate for differences
          ## in theta in the multigroup.data.gen() function
          temp_data <- multigroup.data.gen(n = temp_n, k = temp_k, g = temp_g, 
                                           method = "theta", model = "2PL")
            
          temp_model <- ltm(temp_data[ , 1:temp_k] ~ z1, IRT.param = TRUE)
          
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = MG.LR.test(l_0 = l_0, data = temp_data, g = temp_g, 
                                                 k = temp_k, start.val = start.vals), 
                              df = (temp_g - 1) * 2 * temp_k)
 
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)

        }
      
        TP_2PL <- rbind(TP_2PL, c(a, b, d, mean(prop_1)))
      }
      

  }
  
}


TP_2PL <- TP_2PL %>%
  pivot_wider(., names_from = G, values_from = MG.LR) %>%
  .[-1, -3]
  
rownames(TP_2PL) <- NULL
colnames(TP_2PL) <- c("test_length", "sample_size", "LR2", "LR3", "LR4")
TP_2PL
```


### 3.2. 3PL Condition

```{r Theta Power 3PL }
library(ltm)
library(dplyr)
library(mirt)
library(tidyr)

test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 300, 500, 1000, 1500))
n_groups <- as.factor(c(2, 3, 4))

n_sim <- 300

TP_3PL <- data.frame(test_length = NA, sample_size = NA, G = NA, MG.LR = NA)

set.seed(1248)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
    
    for(d in levels(n_groups)){
      
      prop_1 <- rep(NA, n_sim)
      
      temp_k <- as.numeric(as.character(a))
      temp_g <- as.numeric(as.character(d))
      temp_n <- round(as.numeric(as.character(b))/temp_g)
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){

          ## Now, data generation according to differences in theta and 3PL
          temp_data <- multigroup.data.gen(n = temp_n, k = temp_k, g = temp_g, 
                                           method = "theta", model = "3PL")
            
          temp_model <- ltm(temp_data[ , 1:temp_k] ~ z1, IRT.param = TRUE)
          
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = MG.LR.test(l_0 = l_0, data = temp_data, g = temp_g, 
                                                 k = temp_k, start.val = start.vals), 
                              df = (temp_g - 1) * 2 * temp_k)
 
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)

        }
      
        TP_3PL <- rbind(TP_3PL, c(a, b, d, mean(prop_1)))
      }
      

  }
  
}


TP_3PL <- TP_3PL %>%
  pivot_wider(., names_from = G, values_from = MG.LR) %>%
  .[-1, -3]

rownames(TP_3PL) <- NULL
colnames(TP_3PL) <- c("test_length", "sample_size", "LR2", "LR3", "LR4")

TP_3PL
```


## 4. Power Analysis for differences in item parameters alpha, beta and gamma

### 4.1. 3PL Condition 

```{r Item Param Power 3PL (incl. gamma) }
library(ltm)
library(dplyr)
library(mirt)
library(tidyr)

test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 300, 500, 1000, 1500))
n_groups <- as.factor(c(2, 3, 4))

n_sim <- 300

IPP_a_b_g_3PL <- data.frame(test_length = NA, sample_size = NA, G = NA, MG.LR = NA)

set.seed(1248)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
    
    for(d in levels(n_groups)){
      
      prop_1 <- rep(NA, n_sim)
      
      temp_k <- as.numeric(as.character(a))
      temp_g <- as.numeric(as.character(d))
      temp_n <- round(as.numeric(as.character(b))/temp_g)
      
      start.vals <- start.values(temp_k)
      
        for(i in 1:n_sim){

          ## Now, data generation according to a new function which also
          ## allows for differences in the pseudo-guessing parameter for the 3PL
          temp_data <- gamma.diff.data.gen(n = temp_n, k = temp_k, g = temp_g, 
                                           method = "param", model = "3PL")
            
          temp_model <- ltm(temp_data[ , 1:temp_k] ~ z1, IRT.param = TRUE)
          7
          l_0 <- temp_model$log.Lik
        
          p_value_1 <- 1 - pchisq(q = MG.LR.test(l_0 = l_0, data = temp_data, g = temp_g, 
                                                 k = temp_k, start.val = start.vals), 
                              df = (temp_g - 1) * 2 * temp_k)
 
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)

        }
      
        IPP_a_b_g_3PL <- rbind(IPP_a_b_g_3PL, c(a, b, d, mean(prop_1)))
      }
      

  }
  
}


IPP_a_b_g_3PL <- IPP_a_b_g_3PL %>%
  pivot_wider(., names_from = G, values_from = MG.LR) %>%
  .[-1, -3]

rownames(IPP_a_b_g_3PL) <- NULL
colnames(IPP_a_b_g_3PL) <- c("test_length", "sample_size", "LR2", "LR3", "LR4")

IPP_a_b_g_3PL
```


## 5. Session Information for Reproducibility

```{r session info }
sessionInfo()
```

