---
title: "Master Thesis Code"
author: "Nina van Gerwen (1860852)"
date: "2022-10-28"
output: html_document
---

## Functions for IRF calculations

These are necessary in data generation as we will vary the data generating model.

```{r IRF calculation functions}
one.pl <- function(theta, beta){
  prob <- exp(theta - beta) / (1 + exp(theta - beta))
  return(prob)
}
## Question Dave: is er een probleem met ene functie - beta en andere + beta

two.pl <- function(theta, alpha, beta){
  prob <- exp((alpha * theta) + beta) / (1 + exp((alpha * theta) + beta))
  return(prob)
}

three.pl <- function(theta, alpha, beta, gamma){
  prob <- gamma + (1 - gamma) * (exp((alpha * theta) + beta) / (1 + exp((alpha + theta) + beta)))
  return(prob)
}

```

## Function for data generation

Now we create the function for data generation, where the input is 
sample size, test length and the true model.

```{r Data generation functions}
data.gen <- function(n, k, model = "1PL"){
  theta <- matrix(data = rep(rnorm(n), k), ncol = k)
  
  beta <- matrix(data = rep(c(-2, -1, 0, 1, 2), k), nrow = n, ncol = k, byrow = TRUE)
  
  if(model == "1PL"){
    Z <- one.pl(theta = theta, beta = beta)
  
  }
  if(model == "2PL"){
      
    alpha <- matrix(data = rep(c(0.7, 0.85, 1, 1.15, 1.3), k), ncol = k,
                    nrow = n, byrow = TRUE)
  
    Z <- two.pl(theta = theta, alpha = alpha, beta = beta)
    
  }
  if(model == "3PL"){
    
    alpha <- matrix(data = rep(c(0.7, 0.85, 1, 1.15, 1.3), k), ncol = k,
                    nrow = n, byrow = TRUE)
  
    gamma <- matrix(data = rep(c(0.05, 0.08, 0.11, 0.14, 0.17), k), ncol = k, 
                    nrow = n, byrow = TRUE)
  
    Z <- three.pl(theta = theta, alpha = alpha, beta = beta, gamma = gamma)
  }
  
data <- matrix(data = rbinom(n = n * k, size = 1, prob = Z), ncol = k, nrow = n)
  
return(data)

}
```

## Functions for model fit calculations

To fit the 1PL model, we can use the rasch() function.
To fit the 2PL model, we can use the ltm() function.
To fit the 3PL model, we can use the tpm() function.

```{r model fit calculation functions}

LR.test <- function(data, model = "1PL", g = 2){

  temp_data <- as.data.frame(data)
  temp_data$split <- sample(1:g, nrow(data), replace = TRUE)

  if(model == "1PL"){
    whole_log <- rasch(data = data)$log.Lik
  }
  if(model == "2PL"){
    whole_log <- ltm(data ~ z1)$log.Lik
  }
  if(model == "3PL"){
    ## safeguard voor convergeren
    whole_log <- tpm(data, start.val = "random")$log.Lik
  }
  
  split_LR_values <- rep(NA, g)
  for(j in 1:g) {
    if(model == "1PL"){
      
      
      split_LR_values[j] <- rasch(subset(temp_data, split == j, select = -split))$log.Lik
    }
    
    if(model == "2PL"){
      
      
      split_LR_values[j] <- ltm(subset(temp_data, split == j, select = -split) ~ z1, IRT.param = TRUE)$log.Lik
    }
    if(model == "3PL"){
      
      
      split_LR_values[j] <- tpm(subset(temp_data, split == j, select = -split),
                                start.val = "random")$log.Lik
    }
  }

  LR_value <- whole_log / prod(split_LR_values, na.rm = TRUE)
  return(LR_value)
}

```

voor calculatie van de pchisq moet het 1 - pchisq() zijn, en we hebben
functie nodig voor calculeren van DF.

```{r degree of freedom function}
degree.function <- function(model = "1PL", k, g = NULL){
  if(is.null(g)){
    if(model == "1PL"){
      df <- k
    }
    if(model == "2PL"){
      df <- 2 * k
    }
    if(model == "3PL"){
      df <- 3 * k
    }
    
  } else {
    if(model == "1PL"){
      df <- g * k
    }
    if(model == "2PL"){
      df <- 2 * g * k
    }
    if(model == "3PL"){
      df <- 3 * g * k
    }
    
  }
  return(df)
}

```


BOVENDIEN:

als je een model dat genest is onder een ander model test, dan wil je niet
dat het verworpen wordt (je test nog steeds alpha).

VOORBEELD:

Power is als: data onder 3PL gecreerd, en je test het 1PL/2PL model.

Alpha is als: data onder 1PL gecreerd, en je test het 1/2/3PL.

## Functions for TFI/CLI calculatoins

```{r TLI/CLI functions}

```

CONVERGENCE IS EEN GROOT PROBLEEM. Wellicht een eigen 'control' opzetten
(zie ?tpm)? Wellicht deze hele these weggooien?

## Testing the functions

```{r}

space <- list()

for(i in 1:100){
  temp_data <- data.gen(100, 5, "1PL")
  
  
}

library(ltm)
test_dat <- data.gen(50, 15, "3PL")
mml1 <- rasch(test_dat)
mml2 <- ltm(test_dat ~ z1, IRT.param = TRUE)
mml3 <- tpm(test_dat, start.val = "random")

anova(mml1, mml2)
anova(mml2, mml3)
```

```{r Final run}
test_length <- as.factor(c(5, 10, 20))
sample_size <- as.factor(c(200, 500, 1000, 1500, 2000))
true_model <- as.factor(c("1PL", "2PL", "3PL"))

n_sim <- 5

## We vary three factors, and for the simulation study, we 
## cross-examine them through a nested for loop

## TUPLES

results <- data.frame(test_length = NA, sample_size = NA,
                      true_model = NA, LR2 = NA,
                      LR3 = NA, LR4 = NA, Chisq = NA, P_Chisq = NA)

for(a in levels(test_length)){

  for(b in levels(sample_size)){
      
    for(c in levels(true_model)){
      
      
      prop_1 <- rep(NA, n_sim)
      prop_2 <- rep(NA, n_sim)
      prop_3 <- rep(NA, n_sim)
      prop_4 <- rep(NA, n_sim)
      prop_5 <- rep(NA, n_sim)
      
        for(i in 1:n_sim){
        ## Then, for every condition, generate data according to the
        ## current condition
          temp_data <- data.gen(n = as.numeric(as.character(b)), 
                                k = as.numeric(as.character(a)), 
                                model = as.character(c))
        
        
          p_value_1 <- 1 - pchisq(q = -2 * LR.test(data = temp_data, 
                                              model = "3PL", g = 2), 
                   df = 1)
 
          p_value_2 <- 1 - pchisq(q = -2 * LR.test(data = temp_data, 
                                              model = "3PL", g = 3),
                              df = 1)
        
          p_value_3 <- 1 - pchisq(q = -2 * LR.test(data = temp_data, 
                                              model = "3PL", g = 4), 
                              df = 1)
          
          p_value_4 <- 1
          
          p_value_5 <- 1
          
        
          prop_1[i] <- ifelse(p_value_1 < .05, 1, 0)
          prop_2[i] <- ifelse(p_value_2 < .05, 1, 0)
          prop_3[i] <- ifelse(p_value_3 < .05, 1, 0)
          prop_4[i] <- ifelse(p_value_4 < .05, 1, 0)
          prop_5[i] <- ifelse(p_value_5 < .05, 1, 0)
        }
      
      results <- rbind(results, c(a, b, c, mean(prop_1), mean(prop_2), 
                                  mean(prop_3), mean(prop_4), mean(prop_5)))

    }
  }
  
}

results <- results[-1 ,]
rownames(results) <- NULL

```

      

      

        
        

